05 - PubliliaFitness - Estimating lifetime fecundity and fitness

### 2023-08-27 scripts/Publilia-01-qc.Rmd
Starting script 05 from `strats_data` instead of `visited_1st_4d_data`

## `Publilia-04-within_clutch_iteroparity.Rmd`
```{r}
write_rds(clutch_egg_num_data, here("saved_rds/clutch_egg_num_data"))
write_rds(t5eggs_data, here("saved_rds/t5eggs_data"))
write_rds(fec_model_data, here("saved_rds/fec_model_data.rds"))
write_rds(clutch_status_init_hatch_labels_v1, here("saved_rds/clutch_status_init_hatch_labels_v1.rds"))
```

>>>>

## `Publilia-04-within_clutch_iteroparity.Rmd`
```{r}
write_rds(clutch_egg_num_data, here("saved_rds/clutch_egg_num_data.rds"))
write_rds(t5eggs_data, here("saved_rds/t5eggs_data"))
write_rds(fec_model_data, here("saved_rds/fec_model_data.rds"))
write_rds(clutch_status_init_hatch_labels_v1, here("saved_rds/clutch_status_init_hatch_labels_v1.rds"))
```

# 14 - Write data for

```{r}
write_rds(moms_plant_data, here("saved_rds/moms_plant_data.rds"))
write_rds(focalMb4h_sc2h, here("saved_rds/focalMb4h_sc2h.rds"))
write_rds(clutch_survivorship_data, here("saved_rds/clutch_survivorship_data.rds"))
```

### 2023-08-27 scripts/Publilia-05-fitness.Rmd
Starting script 05 from `strats_data` instead of `visited_1st_4d_data`

```{r}
library(multcompView)
source(here("R/Publilia-05-fitness-functions.R"))
```

@@

```{r}
surv_reg_data <- read_rds(here("saved_rds/surv_reg_data.rds"))
surv_model_fits <- read_rds(here("saved_rds/surv_model_fits.rds"))
```

@@

```{r}
est_fitness_prep <- visited_1st_4d_data %>%
  filter(!is.na(initiation_date)) %>%
  group_by(Female_ID) %>%
  mutate(started_by_initiating =
           any(clutch_visited_order==1&clutch_initiated_order==1)) %>%
  ungroup() %>%
  select(location, Female_ID, clutch_visited_order, clutch_initiated_order, 
         desert_rel2h = est_rel2h, guard_duration,
         mom_at_init, started_by_initiating) %>%
  left_join(select(clutch_survivorship_data, location:last_desert, 
                   n_eggs, med_t5eggs, n_hatched, n_survived, DormPlant), 
            by=c("location")) %>%
  left_join(select(focalMb4h_sc2h, Female_ID, location, focalMb4h_sc2h),
            by=c("Female_ID", "location")) %>%
  mutate(des=last_des_rel2h, 
         ants=ifelse(ants_removed,"Removed","Control"),
         vis=`visited in 4 days`) %>%
  left_join(select(ant_counts, location, ever_antsP)) %>%
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  left_join(dplyr::select(moms_plant_data, DormPlant, last_visit_ends_P), by="DormPlant") %>%
  mutate(last_desP_rel2h = last_visit_ends_P-est_hatch,
         ant_treatment=ifelse(ants_removed,"Removed","Control"))

@@

# It appears that my modeling is overestimating the number of nymphs hatched for clutches with very low est hatched v1
est_fitness_clutch %>% 
  filter(!is.na(est_hatched_clutch_v1),!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=est_hatched_clutch_v1, y=est_hatched_clutch_v6, color=sum_vD==0)) +
  geom_abline(slope=1) +
  geom_point()

@@

split violin v6
```{r}
est_fitness_life_strats_splitvio_eggs_v6.p <- est_fitness_life_strats %>%
  mutate(desert_cat = fct_relevel(desert_cat, "visit", "desert", "eggs", "nymphs")) %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed, color=ants_removed)) +
  geom_split_violin(color=NA, alpha=0.50) +
  geom_point(position=position_jitterdodge(jitter.width=0.08, dodge.width=0.3), pch = 16, size=3) +
  ylab("Estimated lifetime fecundity\n(number of eggs laid)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "black")) +
  scale_color_manual(labels=c("Control","Ants removed"), values=c("brown1", "black")) +
  theme_publilia()

>>>>

```{r}
library(here)
library(tidyverse)
library(rstatix)
library(effectsize)
library(multcompView)
source(here("R/utils.R"))
source(here("R/Publilia-05-fitness-functions.R"))
```

@@

```{r}
strats_data <- read_rds(here("saved_rds/strats_data.rds"))
clutch_survivorship_data <- read_rds(here("saved_rds/clutch_survivorship_data.rds"))
focalMb4h_sc2h <- read_rds(here("saved_rds/focalMb4h_sc2h.rds"))
moms_plant_data <- read_rds(here("saved_rds/moms_plant_data.rds"))
surv_model_fits <- read_rds(here("saved_rds/surv_model_fits.rds"))
```

@@

```{r}
est_fitness_prep <- strats_data %>%
  filter(!is.na(initiation_date)) %>%
  group_by(Female_ID) %>%
  mutate(started_by_initiating =
           any(clutch_visited_order==1&clutch_initiated_order==1)) %>%
  ungroup() %>%
  select(location, Female_ID, clutch_visited_order, clutch_initiated_order,
         desert_rel2h = est_rel2h, guard_duration,
         mom_at_init, started_by_initiating, ever_antsP) %>%
  left_join(select(clutch_survivorship_data, location:last_desert,
                   n_eggs, med_t5eggs, n_hatched, n_survived, DormPlant),
            by=c("location")) %>%
  left_join(select(focalMb4h_sc2h, Female_ID, location, focalMb4h_sc2h),
            by=c("Female_ID", "location")) %>%
  mutate(des=last_des_rel2h,
         ants=ifelse(ants_removed,"Removed","Control"),
         vis=`visited in 4 days`) %>%
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  left_join(dplyr::select(moms_plant_data, DormPlant, last_visit_ends_P), by="DormPlant") %>%
  mutate(last_desP_rel2h = last_visit_ends_P-est_hatch,
         ant_treatment=ifelse(ants_removed,"Removed","Control"))

# beware that clutch_status_momDays_b4h was create by right-joining clutch_momDays_b4h; possible that some hidden filtering occurred there

@@

# It appears that my modeling is overestimating the number of nymphs hatched for clutches with very low est hatched v1
est_fitness_clutch %>% 
  filter(!is.na(est_hatched_clutch_v1),!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=est_hatched_clutch_v1, y=est_hatched_clutch_v6, color=vis)) +
  geom_abline(slope=1) +
  geom_point()

@@

split violin v6
```{r, fig.width = 6, fig.height = 4}
est_fitness_life_strats_splitvio_eggs_v6.p <- est_fitness_life_strats %>%
  mutate(desert_cat = fct_relevel(desert_cat, "visit", "desert", "eggs", "nymphs")) %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed, color=ants_removed)) +
  geom_split_violin(color=NA, alpha=0.50) +
  geom_point(position=position_jitterdodge(jitter.width=0.08, dodge.width=0.3), pch = 16, size=3) +
  ylab("Estimated lifetime fecundity\n(number of eggs laid)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "black")) +
  scale_color_manual(labels=c("Control","Ants removed"), values=c("brown1", "black")) +
  theme_publilia()
